The emotion--facial expression link: Evidence from human and automatic expression recognition

### 2020-11-21 manuscript/manuscript_apa.tex
final

The constructivist approach claims that facial displays are behaviors whose meaning is inferred by perceivers. Findings support this perceiver dependence (Lindquist \& Gendron, 2013; Niedenthal et al., 2017). They show that to make meaning of another person's facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories. For instance, Gendron and her colleagues used a face-sorting task allowing them to manipulate the influence of emotion concepts on how facial expressions were perceived. They conducted their experiment among U.S. participants and Himba participants from remote regions of Namibia (Gendron et al., 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron et al., 2018) both groups with limited exposure to Western culture. Gendron and her colleagues demonstrated that facial expressions were not universally recognized in discrete emotional terms. Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn't have a necessary relationship to emotions. They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors (see also Crivelli et al., 2017 for similar observations among a small-scale society of Papua New Guinea). The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are cast on the perceived face to make meaning of the expresser's facial displays. Following this approach, faces convey a range of information essential for social communication. They are best conceived as tools displaying signals in social interactions (Crivelli \& Fridlund, 2018). These signals can convey individuals' motivations and readiness (Frijda \& Tcherkassof, 1997) or social messages (Fridlund, 1994). As for emotional meaning, more specifically, this is shaped by the perceiver according to the specific context in which the facial displays are observed.

@@

\leavevmode\hypertarget{ref-gendron2018emotion}{}%
Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). \emph{Emotion perception in hadza hunter gatherers}. \url{https://doi.org/10.31234/osf.io/pf2q3}

>>>>

The constructivist approach claims that facial displays are behaviors whose meaning is inferred by perceivers. Findings support this perceiver dependence (Lindquist \& Gendron, 2013; Niedenthal et al., 2017). They show that to make meaning of another person's facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories. For instance, Gendron and her colleagues used a face-sorting task allowing them to manipulate the influence of emotion concepts on how facial expressions were perceived. They conducted their experiment among U.S. participants and Himba participants from remote regions of Namibia (Gendron et al., 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron et al., 2020) both groups with limited exposure to Western culture. Gendron and her colleagues demonstrated that facial expressions were not universally recognized in discrete emotional terms. Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn't have a necessary relationship to emotions. They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors (see also Crivelli et al., 2017 for similar observations among a small-scale society of Papua New Guinea). The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are cast on the perceived face to make meaning of the expresser's facial displays. Following this approach, faces convey a range of information essential for social communication. They are best conceived as tools displaying signals in social interactions (Crivelli \& Fridlund, 2018). These signals can convey individuals' motivations and readiness (Frijda \& Tcherkassof, 1997) or social messages (Fridlund, 1994). As for emotional meaning, more specifically, this is shaped by the perceiver according to the specific context in which the facial displays are observed.

@@

\leavevmode\hypertarget{ref-gendron2018emotion}{}%
Gendron, M., Hoemann, K., Crittenden, A. N., Mangola, S. M., Ruark, G. A., \& Barrett, L. F. (2020). Emotion perception in hadza hunter-gatherers. \emph{Scientific Reports}, \emph{10}(1), 1--17. \url{https://doi.org/10.1038/s41598-020-60257-2}
